Speakers of diverse languages structure their utterances for efficient communication

### 2019-10-02 Paper/info-curves.Rmd
introduction edits and smoothing in methods added to paper

author: 
  - name          : "Josef Klafka"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "5000 Forbes Ave"
    email         : "klafka@cmu.edu"
  - name          : "Daniel Yurovksy"
    affiliation   : "1,2"

@@

One of the defining features of human language is its power to transmit information. We use language for a variety of different proximal tasks tasks: greeting friends, taking notes, signalling group identities. But all share a common unifying purpose: changing the mental state of the listener or reader of our communications [@austin1975]. Language can thus naturally thought of as a code that allows speakers to turn their intended meanings into a form that can be transmitted through the air (or paper, or electrons) and subsequently converted by listeners back into an approximation of the intended meaning [@shannon1948].

Beyond its utility as a metaphor, this coding perspective on language is powerful because it affords a framework for rational analysis: If language has evolved to be an optimal code for information transmission, how should this code be structured [@anderson1989]? The optimal code would have to optimize two competing pressures: (1) a pressure for listeners to easily and successfully decode messages sent by the speaker, and (2) a pressure for speakers to easily code their messages and transmit them with minimal effort and error. A fundamental constriaint on both of these processes is the serial nature of human language: sounds are produced one at a time and each is unavailable perceptually once it is no longer being produced.

The solution that languages have arrived on is incremental processing: Listeners process speech continuously as it arrives, predicting upcoming words and building expectations about the likely meaning of utterances in real-time rather than at their conclusion [@kutas2011; @tanenhaus1995; @pickering2013]. Because prediction errors can lead to severe processing costs and difficulty integrating new information, speakers should seek to minimize prediction errors. However, the cost of producing more predictable utterances is using more words. Thus, an optimal coding strategy is for speakers seeking to minimize their production costs is to produce utterances that are are just at the prediction capacity of listeners without exceeding this capacity [@aylett2004; @genzel2002].

Using information theory--a mathematical framework for formalizing predictability--researchers have tested and confirmed this general prediction of optimal coding across levels and contexts of language production. For instance, @genzel2002 provided a clever indirect test of this hypothesis at the level of sentences. They showed that the predictability of successive sentences in a discourse--analyzed in isolation--decreases, as would be expected if readers use prior sentences to predict the content of future sentences and total predictability remains constant.

the predicted if the amount of information is constant readers use prior sentences in a discourse to predict the content of future setences as they read.

This prediction of optimal coding--that speakers should maintain a constant rate of predictability has been tested and confirmed in a variety of ways across levels of language.

@@

Due to their differences in distributional frequency, words vary in how much information content they contain [@shannon1948]. Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be [@zipf1949]. Over and above frequency, contextual information content predicts word length, due to speakers minimizing effort in the amount of information they encode in speech [@piantadosi2011]. Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side [@jaeger2007]. Across utterances, context allows speakers to produce more informative and surprising sentences later in communication. The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains [@genzel2002]. From the level of individual words to sentences in paragraphs, information content shapes and is shaped by the context it appears in.

The pressure for efficient communication influences language at all timescales, from online speech production to evolution. @mahowald2013 found that speakers tend to use shorter, more efficient word forms in predictive contexts, minimizing communicative cost without sacrificing intelligibility. Predictability and ambiguity shape the distribution of color terms in languages [@gibson2017] as well as word lengths [@piantadosi2011] across languages, and speakers are able to deterministically decode words despite ambiguity in word forms [@piantadosi2012].

How do frequency and predictive context determine how we distribution information at the level of utterances? We do not speak only in phrases or paragraphs, but conversationally we speak in utterances and in writing we produce sentences as a coherent unit of a complete thought. Languages possess very different grammatical and morphological structures--the same thought can be encoded in at least as many ways as there are different languages. The average rate of information speakers transfer in syllables differs from language to language [@aylett2004]. How do the different structures seen in different languages, with different rates of information encoding, affect how speakers structure information in production?

@yu2016 proposed a novel method for quantifying the amount of information typically found at each part of English utterances. Treating each word position by itself, @yu2016 calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty [@shannon1948]. They found a distinctive three-step shape to the distribution of information in English written sentences.

In this paper, we expand on this body of prior work in a number of novel ways. We replicate the results from @yu2016 with a different metric, one intimately tied to incremental word processing. We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech. We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate past the initial words. We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information. We find that each language has a characteristic, robust frequency-based information distribution tied to the language's morphology, phonology and syntax. Speakers of a language will tend to distribute information according to that language’s distribution of information in their utterances. As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.

@@

>>>>

author: 
  - name          : "Josef Klafka"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "5000 Forbes Ave"
    email         : "jklafka@andrew.cmu.edu"
  - name          : "Daniel Yurovsky"
    affiliation   : "1,2"

@@

One of the defining features of human language is its power to transmit information. We use language for a variety of different tasks: greeting friends, taking notes, signaling group identities. But all share a common unifying purpose: changing the mental state of the listener or reader of our communications [@austin1975]. Language can thus naturally thought of as a code that allows speakers to turn their intended meanings into a message that can be transmitted through the air (or paper, or electrons) and subsequently converted by listeners back into an approximation of the intended meaning [@shannon1948].

Beyond its utility as a metaphor, this coding perspective on language is powerful because it allows a framework for rational analysis: If language has evolved to be an optimal code for information transmission, how should this code be structured [@anderson1989]? The optimal code would have to work with two competing pressures: (1) a pressure for listeners to easily and successfully decode messages sent by the speaker, and (2) a pressure for speakers to easily code their messages and transmit them with minimal effort and error. A fundamental constriaint on both of these processes is the linear order of spoken language: sounds are produced one at a time and each is unavailable perceptually once it is no longer being produced.

The strategic solution people employ is *incremental processing*. Listeners process speech continuously as it arrives, predicting upcoming words and building expectations about the likely meaning of utterances in real-time rather than at their conclusion [@kutas2011; @tanenhaus1995; @pickering2013]. Because prediction errors can lead to severe processing costs and difficulty integrating new information, speakers should seek to minimize prediction errors. However, the cost of producing more predictable utterances is using more words. Thus, an optimal coding strategy is for speakers seeking to minimize their production costs is to produce utterances that are are just at the prediction capacity of listeners without exceeding this capacity [@aylett2004; @genzel2002].

Using information theory--a mathematical framework for formalizing predictability--researchers have tested and confirmed this general prediction of optimal coding across several levels and contexts of language production. For instance, @genzel2002 provided a clever indirect test of this hypothesis across sentences. They showed that the predictability of successive sentences in a discourse--analyzed in isolation--decreases, as would be expected if readers use prior sentences to predict the content of future sentences and total predictability remains constant.

This prediction of optimal coding--that speakers should maintain a constant rate of predictability has been tested and confirmed in a variety of ways across levels of language. At the level of individual words, @mahowald2013 showed that speakers use shorter words in more predictive contexts and prefer the shorter versions of words in more predictive contexts, maximizing the amount of information in each word while minimizing the time spent on those words. Over time, @piantadosi2011 showed that shorter words across languages tend to be more easily predictable in context and thus less informative, over and above frequency.

Efficient encoding also impacts how speakers structure phrases. The inclusion of complementizers in relative clauses [@jaeger2007] and the use of contractions [@frank2008] are two situations in sentence formation in which speakers use structure to communicate efficiently, spreading more information across a larger amount of linguistic material.

How these pressures shape language evolution??

However, despite this literature using the predictive coding model of language, one level has not yet been studied in depth: how speakers structure individual utterances. This level may show the strongest effects of variation between languages, as specific languages have properties that constrain how speakers may form utterances in those languages, such as canonical word order. These properties vary widely from language to language.

@yu2016 studied the utterance level in written English sentences using a contextless entropy model based on word frequency. They found a distinctive three-step distribution regardless of sentence length, with little information in the first words of sentences and the most information in the final words. This was surprising, as the distribution they found was robustly different from the linearly increasing trend in sentences from @genzel2002, and also did not resemble the uniform distribution of information that one might expect from a communicative efficiency account, in which each word has approximately equal information close to the channel capacity.

In this paper, we expand on this body of prior work in a number of novel ways. We replicate the results from @yu2016 with a metric tied to incremental word processing. We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech. We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate past the initial words. We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information. We find that each language has a characteristic, robust frequency-based information distribution tied to the language's morphology, phonology and syntax. Speakers of a language will tend to distribute information according to that language’s distribution of information in their utterances. As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.

@@

@@

In our contextual models, we face the problem of overfitting: we only train our model on those utterances which occur in the corpus and test our model on the same utterances. This ignores possible other utterances which the speakers could have produced, e.g. the words "I", "saw" and "bears" are in the corpus vocabulary, which the speaker may not have produced in the corpus but could have produced. To combat this issue, we use modified Kneser-Ney smoothing [@chen1999empirical].

In this smoothing technique, we filter the ngram counts from the corpus to only those above a threshold, ignoring rare ngrams. We then apply *absolute discounting* to all of the ngram, which reduces the counts of all observed ngrams in the corpus by a fixed number, reducing the impact of uncommon ngrams on our surprisal calculations. We *interpolate* lower order ngram counts when computing bigrams and trigrams: this means computing a weighted average of the ngram counts along with lower order ngram counts, in case the higher order counts for a particular ngram are rare. The main insight of Kneser-Ney smoothing is to weight ngram counts by how often the final word occurs as a continuation of any previous words in the corpus, downweighting common unigrams which occur only in a few contexts. This version of modified Kneser-Ney smoothing is in the Kenlm language modeling toolkit [@heafield2013scalable].
